Geometric variational problems

### 2020-11-10 NSF2020/Description/specificaim2_ver2.tex
RJR

The hydrodynamic interactions enter through the transmission of viscous stresses from one particle to the another.  
We must solve the system 
\setcounter{midequation}{\theequation}
\addtocounter{midequation}{2}

@@

\noindent for $i = 1,\dots, N$.
The inputs to \eqref{St1}--\eqref{HSB2} are the collection of particle configurations.
Within this system, the activity $u$ solves the screened Laplace equation \eqref{SL1} with the material label boundary condition \eqref{SL2}
and far-field condition \eqref{SL3}. The forces ${\bf F}_i$ and torques ${\bf G}_i$ come from \eqref{totalFG}.
In turn, the fluid velocity ${\bf u}$ solves the Stokes equations (\ref{St1}, \ref{St2}) with background flow boundary condition (\ref{St3}).
The force and torque equations \eqref{HSB1}, \eqref{HSB2} couple the (\ref{St1}--\ref{St4}) to (\ref{SL1}--\ref{SL3}).
The outputs are the collection of velocity and rotation pairs $({\bf v}_i, {\bf w}_i)$ that uniquely satisfy the rigid body boundary condition \eqref{St3}.

@@

We simulate two-dimensional vesicle bilayers in external flows using \eqref{St1}--\eqref{HSB2}.
The vesicle bilayer is formed by a many-body collection of amphiphilic particles. There are two leaflets: 
the inner leaflet consisting of particles in contact and pointing their hydrophobic tails away from the vesicle interior;
the particles in the outer leaflet are in contact with the surrounding fluid and point their tails inward.
The particle laden flow changes the shape of the bilayer, and in return the hydrophobic stresses coming from the altered shape 
impart a force on the flow. In other words, there is a two-way coupling between the flow and bilayer shape.

@@

The background flows allow us to study further properties of amphiphilic particle collections. 
Area conservation
Volumne conservation
Analyzing the dynamics compared to previous works
Interaction between the fluid and membrane

We prove a, to our knowledge new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
\begin{equation}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}
As described in Specific Aim 3, we solve \eqref{SL1}--\eqref{SL3}  using a double layer potential representation.
While accurate in the far-field, this representation involves a strongly singular integral which
leads to large numerical errors when evaluating the the hydrophobic stress \eqref{stress} along the particle boundaries.
The identity in Proposition \ref{prop:recip} gets around this critical problem by utilizing the fact that $u_j$, $j\neq i$ is smooth and the normal derivative
of  $u_i$ continuous across $\partial P_i$.

>>>>

\textbf{Model equations.}
The hydrodynamic interactions enter through the transmission of viscous stresses from one particle to the another.  
We must solve the system 
\setcounter{midequation}{\theequation}
\addtocounter{midequation}{2}

@@

\noindent for $i = 1,\dots, N$.
The equations \eqref{St1}--\eqref{HSB2} express a two-way coupling between the flow and bilayer shape.
The particle laden flow changes the shape of the bilayer, and in return the hydrophobic stresses coming
from the altered shape impart a force on the flow. 
The inputs to \eqref{St1}--\eqref{HSB2} are the collection of particle configurations.
Within this system, the activity $u$ solves the screened Laplace equation \eqref{SL1} with the material label boundary condition \eqref{SL2}
and far-field condition \eqref{SL3}. The forces ${\bf F}_i$ and torques ${\bf G}_i$ come from \eqref{totalFG}.
In turn, the fluid velocity ${\bf u}$ solves the Stokes equations (\ref{St1}, \ref{St2}) with background flow boundary condition (\ref{St3}).
The force and torque equations \eqref{HSB1}, \eqref{HSB2} couple the (\ref{St1}--\ref{St4}) to (\ref{SL1}--\ref{SL3}).
The outputs are the collection of velocity and rotation pairs $({\bf v}_i, {\bf w}_i)$ that uniquely satisfy the rigid body boundary condition \eqref{St3}.

@@

\textbf{Simulation studies.}
We simulate two-dimensional vesicle bilayers in external flows using \eqref{St1}--\eqref{HSB2}.
The vesicle bilayer is formed by a many-body collection of amphiphilic particles.
At large length scales and for large particle-number, the bilayer behaves as an elastic membrane,
in two-dimensions an elastic curve. 
This will lead to a deeper understanding of the elastic properties established by Specific Aim 2.

@@

We will compare the behavior of our particle-based vesicles in Stokes flow to well-established results.
To test area incompressibility, we form the bilayer midplane by fitting to the particle centers. Under moderate shear rates,
the vesicle elongates and takes on an elliptical shape. We will study the increase in midplane area as a function of shear rate.
Additionally, we will test for tangential divergence of velocity.
Researchers have devised to model area incompressible and
volume conserving membranes
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}.
Our preliminary results show that the particle-based bilayers have the same area modulus, about 60 \kBT,  as that of real lipid bilayer membranes,
and so we expect realistic results from our flow simulations with regard to area incompressibility.  It is noteworthy that
area incompressibility is some sense baked into the model through the hydrophobic attraction.

Bilayer membranes have a small permeability to water \cite{323e9a2f0c58487ea82518d7a1f96485},
and so modelers often assume a vesicle membrane that conserves volume. In our particle-based approach, water motion across
the bilayer is limited by the size of the gaps between the particles, which is an artifact of coarse-graining.
Making these gaps smaller comes at the expense of numerical accuracy, and so we will assess if there is a reasonable trade-off
between approximate volume conservation and simulation complexity.

Finally, the particle-based vesicle bilayers have two distinct leaflets.
The inner and outer leaflets consists of the particles in contact with the vesicle interior and 
surrounding fluid respectively. Since there the separation between these layers is roughly fixed,
about 2 nm, certain fluid mechanical effects arise that are not present in zero-thickness membrane models. 
An important effects we observe is that the leaflets slide against one another under shear flow. 
This implies that part of the viscous dissipation in the aqueous phase is enhanced by intermonolayer friction
\cite{SHKULIPA2005823, ShkulipaThesis}. We will characterize the intermonolayer
slip as a function of shear rate, vesicle diameter and particle geometry.

\textbf{Novel reciprocal relations.}
We prove a, to our knowledge new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
\begin{equation}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}
As described in Specific Aim 3, we solve \eqref{SL1}--\eqref{SL3}  using a double layer potential representation.
While accurate in the far-field, this representation involves a strongly singular integral which
leads to large numerical errors when evaluating the the hydrophobic stress \eqref{stress} along the particle boundaries.
The identity in Proposition \ref{prop:recip} gets around this critical problem by utilizing the fact that $u_j$, $j\neq i$ is smooth and the normal derivative
of  $u_i$ continuous across $\partial P_i$.
